---
title: "Untidy data"
author:
- Callum Abbott
- Georgia Zhao
- Panagiotis Maouris
- Xihang Li
output:
  html_document: default
  pdf_document: default
---

## Setup - Load packages

```{r setup, include = FALSE}
library(dplyr)
library(tidyr)
library(tibble)
library(rvest)
library(lubridate)
library(ggplot2)
```

## Task 1 - Lego Sales Data

```{r message = FALSE}
sales = readRDS("data/lego_sales.rds")
```

### Part 1 - Tidying the data


#### Write up

The lego sales dataset has been tidied with the convention that each row represents
the purchases of a single person of a particular lego set. This means that the same 
purchaser is able to occupy multiple rows in our data set if they buy more than 
one lego set with different `Theme`s. Note however that we have chosen to preserve
the `Quantity` variable since we believe this is a tidier representation of the data, 
and hence, if multiple purchases are made but they are of the same `Theme` of lego set, this person
will only occupy a single row with `Quantity` > 1. A further justification for this
decision is that the `Quantity` variable  is still easily accessible, and hence, 
easy to work with (see Q2). 

Note that we have also chosen to give the hobbies variable 3 distinct columns. 
This is because we believe that the elements of a tidy dataframe should not
be subscriptable i.e. matrices, vectors, lists, etc. 

#### Code

```{r task1_tidy}
# --------------------- #
# TASK: Tidy the data.  #
# --------------------- #

lego_sales_tbl = tibble(lego_sales = sales) %>%
  # Unnesting list of lists to row column structure
  unnest_wider(lego_sales) %>%
  unnest_longer(purchases) %>%
  unnest_wider(purchases) %>%
# Creating distinct columns for the hobbies
unnest_wider(hobbies, names_sep = "_") 
```

```{r task1_top}
# Print out the first 10 lines of your tidy data frame below 
print(head(lego_sales_tbl, n = 10L))
```

### Part 2 - Questions

#### Question 1

```{r task1_q1}
# --------------------------------------------------------------- #
# TASK: What are the three most common first names of purchasers? #
# --------------------------------------------------------------- #

common_buyers_df = lego_sales_tbl %>%
  # Getting distinct purchasers
  distinct(first_name, last_name, phone_number) %>%
  # Counting occurrences of first names
  count(first_name, name = 'name_freq') %>%
  arrange(desc(name_freq)) %>%
  # Getting top 3 purchasers (accounting for ties)
  slice_max(name_freq, n = 3)

cat("The three most common first names of purchasers :",
    as.character(unlist(common_buyers_df[1])),sep = "\n")
```

#### Question 2

```{r task1_q2}
# --------------------------------------------------------- #
# TASK: Which Lego theme has made the most money for Lego?  #
# --------------------------------------------------------- #

themes_df = lego_sales_tbl %>%
  group_by(Theme) %>%
  # Creating new summary column with total money spent for each theme
  summarise(total_USPrice = sum(USPrice * Quantity), .groups = "drop") %>%
  arrange(desc(total_USPrice)) %>%
  # Grabbing theme which has made the most money for Lego (accounting for ties)
  slice_max(total_USPrice, n = 1)

cat("The theme has made the most money for lego is:",
    as.character(unlist(themes_df[1])))
```


#### Question 3

```{r task1_q3}
# ----------------------------------------------------------------- #
# TASK: Do men or women buy more Lego sets (per person) on average? #
# ----------------------------------------------------------------- #

female_df = lego_sales_tbl %>%
  filter(gender == "Female")
# Getting total number of sets bought by female purchasers
total_FemaleLegoSets = sum(female_df$Quantity)
# and the total number of female purchasers
total_Female = nrow(distinct(female_df, first_name))

# Repeating for males
male_df = lego_sales_tbl %>%
  filter(gender == "Male")
total_MaleLegoSets = sum(male_df$Quantity)
total_Male = nrow(distinct(male_df, first_name))

# Comparing whether men or woman purchase more Lego sets on average
if ((total_FemaleLegoSets / total_Female) < (total_MaleLegoSets / total_Male)) {
  print("Men buy more Lego sets (per person) on average")
} else if ((total_FemaleLegoSets / total_Female) == (total_MaleLegoSets / total_Male)) {
  print("Men and Women buy equivalent Lego sets (per person) on average.")
} else {
  print("Women buy more Lego sets on average.")
}
```

#### Question 4

```{r task1_q4}
# ---------------------------------------------------------------- #
# TASK: What are the five most popular hobbies of Lego purchasers? #
# ---------------------------------------------------------------- #

hobbies_tbl = tibble(lego_sales = sales) %>%
  # Getting distinct purchasers and hobbies
  unnest_wider(lego_sales) %>%
  distinct(first_name, last_name, phone_number, hobbies) %>%
  select(hobbies) %>%
  # Placing hobbies into distinct rows 
  unnest_longer(hobbies) %>%
  # Removing NA chr values
  filter(!hobbies == "NA") %>%
  count(hobbies, name = 'Freq') %>%
  arrange(desc(Freq)) %>%
  # Grabbing top 5 most popular hobbies (accounting for ties)
  slice_max(Freq, n = 5)

cat("The five most popular hobbies of Lego purchasers are:",
    as.character(unlist(hobbies_tbl[1])), sep = "\n")
```

#### Question 5

```{r task1_q5}
# ---------------------------------------------------------- #
# TASK: Which area code has spent the most money on Legos?   #
# ---------------------------------------------------------- #

areacode_tbl = lego_sales_tbl %>%
  # Selecting relevant variables
  select(phone_number, USPrice, Quantity)%>%
  # Removing NAs from phone_number
  filter(!is.na(phone_number))%>%
  separate(
    col = phone_number,
    # Separating 9 digit phone number into 3-3-3 
    into = c("First3Digits", "Mid3Digits", "Last4Digits"),
    sep = "-"
  )%>%
  # Grouping by area code
  group_by(First3Digits) %>%
  # Calculating money spent by area code
  summarise(total_money = sum(USPrice * Quantity), .groups = "drop")%>%
  arrange(desc(total_money)) %>%
  # Getting top spending area code (accounting for ties)
  slice_max(total_money, n = 1)

cat("The area code has spent the most money on Legos:",
    as.character(unlist(areacode_tbl[1])))
```



## Task 2 - GitHub and dplyr

```{r message = FALSE}
commits = readRDS("data/dplyr_commits.rds")
```

### Part 1 - Structure of a commit

The commits object is a list of length 474 where each element contains another
list of length 11. Each of these element lists represent the data that is collected 
for each commit made to the dplyr package on GitHub.

Examining the list containing the data on a single commit, we observe the data
to be stored in a hierarchical tree structure. The root level of this tree structure
contains information for identifying and finding the commit on GitHub such as
`sha`, `node_id`, `url`, `html_url` and `comments_url`.

In addition to the data listed above, there are additional lists at the root level
of this tree containing meta-data about the commit. These lists include `commit`,
`author`, `committer`, `parents`, `stats` and `files` where the lengths of these
lists range from 1 to 18. Firstly, the `author` and `committer` lists provide 
general information about the author's and commiter's GitHub profile (who are 
usually the same person). The `stats` list contains information on the number of 
lines of code that have been added and deleted across all files that have been 
included in the commit. Meanwhile, the `files` list details the breakdown of the 
changes made to each individual file included in the commit, as well as each 
file's identifying information. The `parents` list contains information on the
commit history of the repository, including which commit is descended from which. 
This will not be useful for our task. Finally, the `commit` list amalgamates some
of the most useful data from the entire commit into a single list with information 
such as the author, committer and commit's identifying information.

### Part 2 - Tidying the data

#### Write up

From the commits object we decided that each row should represent a unique file 
that was modified in a single commit. This led us to extract the following
information to act as columns in our tidy dataset: `sha` - to uniquely identify a 
single commit; `id` - to aid in uniquely identifying authors who have characters
outside the English alphabet in their names; `author_name` - for mapping ids to 
people's names; `date` - to help summarize the general pattern for development of 
dplyr and aid in visualizing the relationship between the commit history and dplyr release
dates on CRAN; and finally `filename` -  to uniquely identify the file changed 
in a single commit.

We believe that the variables listed above are all the information we need
to answer the following 5 questions. Creating more variables and adding more 
data to our tidy dataframe would be simply polluting the data.

#### Code

```{r task2_tidy}
# --------------------- #
# TASK: Tidy the data.  #
# --------------------- #

commits_tbl = tibble(dplyr_commits = commits) %>%
  # Extracting relevant information
  hoist(
    dplyr_commits,
    sha = c("sha"),
    author_name = c("commit", "author", "name"),
    id = c("author", "id"),
    date = c("commit", "author", "date"),
    files = c("files")
) %>%  select(-dplyr_commits) %>% # Dropping unnecessary info
  # Making single row represent a single file and the commit it was changed in
  unnest_longer(files) %>%
  # Getting file names
  hoist(
    files,
    filename = c("filename")
  ) %>% select(-files) # Dropping unnecessary info
```


```{r task2_top}
# Print out the first 10 lines of your tidy data frame below -->
print(head(commits_tbl, n = 10L))
```


### Part 3 - Questions


#### Question 1

```{r task2_q1}
# ---------------------------------------------------------------------- #
# TASK: Who are the top five contributors (in terms of the most commits) #
# to dplyr in 2020?                                                      #
# ---------------------------------------------------------------------- #

# Some authors use different names while they have the same id. Consequently, 
# the commits of each contributor were counted based on their IDs and not their names.

ID_tbl = commits_tbl %>%
  # Getting distinct commit of each contributor 
  distinct(id, sha) %>%
  # Count the contributions of each author and their ids.
  count(id, name = 'NumberCommits') %>%
  arrange(desc(NumberCommits)) %>%
  # Fetching the ids of the top 5 contributors to dplyr (accounting for ties)
  slice_max(NumberCommits, n = 5)

# Creating a tibble to match authors with IDs
author = commits_tbl %>%
select(name = author_name, ID = id) %>%
distinct(name, ID)

# Fetching the names of the top 5 contributors to dyplyr
topconts = c()
for (i in seq_along(ID_tbl$id)){
topconts[i] = author$name[author$ID == ID_tbl$id[i]][1]
}

cat("The top five contributors (in terms of the most commits) are: \n",
    paste(shQuote(topconts), collapse= ", "))
```

#### Question 2


```{r task2_q2}
# -------------------------------------------------------------------------- #
# TASK: Who is the top contributor who is not a current employee of RStudio? #
# -------------------------------------------------------------------------- #

# Web scraper function to collect names of RStudio team members from 
# RStudio website
get_rstudioteam = function(url) {
  # Reading relevant website
  web_stuff = read_html(url)
  # Extracting staff names
  r_ppl = web_stuff %>%
    html_nodes(".teammember .name") %>%
    html_text()
  r_ppl_df = data.frame(r_ppl)
  
  # Removing '\n' from data collected from html code
  name = r_ppl_df$r_ppl
  name = gsub('\n','', name)
  return(data.frame(name))
}
rstudioteam = get_rstudioteam(url = "https://rstudio.com/about/")

# Tibble of names and IDs of RStudio employees
rstudioppl = commits_tbl %>%
  select(author_name, id) %>%
  distinct(author_name, id) %>%
  filter(author_name %in% rstudioteam$name)

# Tibble of contributors not employed by RStudio
notRcommitID_tbl = commits_tbl %>%
  distinct(id, sha) %>%
  count(id, name = 'NumberCommits') %>%
  arrange(desc(NumberCommits)) %>%
  # Excluding contributors who are Rstudio employees
  filter(!(id %in% rstudioppl$id)) %>%
  # Fetching top contributorID who is not a current employee
  slice_max(NumberCommits, n = 1) 

# Some authors used two different names that were not on the Rstudio Team List. 
# As a result, it was preferable to use their IDs instead of their names.
# (eg Romain Francois and Romain François)
topnotRcom = commits_tbl %>%
  select(author_name, id) %>%
  distinct() %>%
  filter(id == notRcommitID_tbl$id) %>%
  select(author_name)

cat("The top contributor who is not a current employee of RStudio is ",paste(shQuote(topnotRcom)))
```
The **commit** with `sha`  **e5730a2b7c27f1857060f570426f459137f8db49** will be 
described. The `author` of this commit is **Hadley Wickham**, Chief Scientist of
the Rstudio Team. His commit was made on 24 June 2020. The `comment` of this commit
was **Resolve Test failure**. In this commit 8 changes have been made, 4 additions 
and 4 deletions. There is 1 `parent` of this commit and 2 `files` have changed 
`"tests/testthat/test-colwise-mutate-errors.txt"` and 
`"tests/testthat/test-colwise-mutate.R"`.\

The `author` has corrected the error message of the function `mutate_all` where 
a space was missing. Moreover, the argument `round` was replaced by the argument
`length`.

#### Question 3

```{r task2_q3}
# -------------------------------------------------------------------------- #
# TASK: Which four files have been modified most often?                      #
# -------------------------------------------------------------------------- #

files_tbl = commits_tbl %>%
  # Extracting relevant variable and sorting
  select(filename) %>%
  count(filename, name = 'commit_freq') %>%
  arrange(desc(commit_freq)) %>%
  # Fetching the most modified files in dplyr (accounting for ties)
  slice_max(commit_freq, n = 4)

cat("The four files that appear in the most number of commits are: \n", 
    paste(shQuote(files_tbl$filename, type="cmd"), collapse=", "))
```


#### Question 4

```{r task2_q4}
# -------------------------------------------------------------------------- #
# TASK: Describe the general pattern for development of dplyr.               #
# -------------------------------------------------------------------------- #

# For arranging dates temporally rather than alphabetically
days_of_week = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")

temporal_tbl = commits_tbl %>%
  # Collapsing tibble to distinct commits
  distinct(sha, date) %>%
  # Extracting time information from data column
  select(date) %>%
  mutate(weekday = weekdays(as.Date(date), abbr=TRUE)) %>%
  mutate(weekday = factor(weekday, levels = days_of_week)) %>%
  # Extracting time from date variable
  separate(
    col = date,
    into = c("date", "time"),
    sep = "T"
  ) %>%
  # Extracting info by hour, minute and time
  separate(
    col = time,
    into = c("hour", "minute", "second"),
    sep = ":"
  ) %>%
  # Removing unnecessary information for task
  select(-minute, -second) %>%
  group_by(weekday, hour) %>%
  count(hour, name = 'freq')

# Plotting
ggplot(
  data = temporal_tbl,
  mapping = aes(
    x = hour,
    y = freq,
    group = 1,
    color = weekday
  )
) + 
geom_line() + 
  # Creating grid that shares weekday variable
  facet_grid(weekday ~ .) + 
  theme_light() + 
  labs(
    title = 'Commits to dplyr by hour',
    x = 'Hour of the Day',
    y = 'Number of Commits',
    color = 'Weekday'
  ) +
  scale_y_continuous(
    limits = c(0, 23),
    n.breaks = 3
  )
```

If we start our analysis by making the reasonable assumption that a larger number
of commits corresponds to more work being done on the dplyr package, we can observe
that Monday and Wednesday seem to be the most productive days of the week
for dplyr contributors respectively.

Now taking on a more holistic view, it seems the majority of commits made
to dplyr occur between the hours 11:00 - 17:00 across all days of the week
which is portraying of the workflow of a typical working day. There is one exception 
however occurring at 08:00 on Mondays. If we postulate that commits are only
counted once those changes have been merged into the master branch, this hump
could be explained by someone at RStudio merging in the changes made by 
people over the weekend every Monday morning.

We also note that no commits were made at 02:00 in the data provided.

Finally, it is also interesting to also observe that a greater quantity of
commits seem to be made on Sundays rather than Saturdays and moreover, that the 
majority of commits made on Sunday seem to occur between the hours of 14:00 and
16:00. Perhaps this could be due to employees at RStudio trying to get ahead
of the work week or maybe the dplyr contributors love coding in R so much they 
can't stop themselves from working!


#### Question 5

```{r task2_q5}
# -------------------------------------------------------------------------- #
# TASK: create a visualization exploring an interesting relationship between #
# these releases and the commit history.                                     #
# -------------------------------------------------------------------------- #

# Getting 2020 releases of dplyr
dplyr_archive = tribble(
       ~ver          ,   ~release_dates    ,
      "0.8.4"	       ,      "2020-01-31" 	 ,	 
      "0.8.5"        ,      "2020-03-07"   , 
      "1.0.0"	       ,      "2020-05-29" 	 ,  
      "1.0.1"	       ,      "2020-07-31" 	 ,
      "1.0.2"        ,      "2020-08-18"   ,
)


# Mutating release_dates into date variable type
dplyr_archive = dplyr_archive %>%
  mutate(release_dates_ymd = as.Date(release_dates)) 


# The dates of the commits made to the dplyr on GitHub in the form Y-M-D  
commitsdates = commits_tbl %>%
  select(sha, date) %>%
  distinct() %>%
  mutate(dateYMD = as.Date(date)) %>%
  select(dateYMD) %>% 
  group_by(dateYMD) %>%
  summarise(NumberCommits = n(), .groups = "drop_last") 

# Plotting the number of commits for each day and the dplyr 
# releases using dotted vertical lines.
ggplot(
  data = commitsdates,
  mapping = aes(
    x = dateYMD,
    y = NumberCommits,
    colour = "Number of Commits",
    )
  ) +
  geom_step() +
  geom_vline(
    xintercept = dplyr_archive$release_dates_ymd, 
    linetype = "dashed", 
    colour = "blue"
  ) + 
  labs(
    title="Commit history and dplyr releases",
    subtitle = "dyplr releases are represented by blue dashed lines",
    color="Labels",
    x = "Months 2020"
  )
```


The graph shows information about the history of commits and the
dates of the 5 dplyr releases in 2020. What comes out from the
graph is that in the period just before the dplyr releases in
2020, the number of commits increased significantly. In addition,
another interesting point is that in the period just after the
dplyr releases the amount of commits went down considerably.
Finally, it is also interesting to notice that a significant number of commits
occur on the day of the package release! This indicates that some of the 
dplyr tests probably didn't encompass all possible edge cases which is very
understandable.
Overall, the authors tend to commit very often when a new dplyr
version is going to be released.

